---
title: 'Example: Bird surveys in central AZ - PHX metro area'
author: "Alessandra Vidal Meza"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
```

Let's load in the data using the **here** package.

```{r}
bird_observations <- read_csv(here::here("data", "52_pp52_birds_1.csv")) %>%
   clean_names()

sites <- read_csv(here::here("data", "52_pp52_sites_1.csv")) %>%
   clean_names()

surveys <- read_csv(here::here("data", "52_pp52_surveys_1.csv")) %>%
   clean_names

taxa_list <- read_csv(here::here("data", "52_pp52_taxalist_1.csv")) %>%
   clean_names()
```

Now, let's create a subset for birds with `species_id` "BHCO" and "RWBL", and from sites with `site_id` "LI-W" and "NU-C".

```{r}
bird_subset <- bird_observations %>%
  filter(species_id %in% c("BHCO", "RWBL") & 
           site_id %in% c("LI-W", "NU-C"))
```

First, left join the **bird_subset** to **sites** by the `site_id` and to **taxa_list** by `species_id`:

```{r}
bird_left <- left_join(bird_subset, sites, by = "site_id") %>%
  left_join(bird_subset, taxa_list, by = "species_id" )
```

Now, a full join. Note that a `full_join()` between **bird_subset** and **sites** will contain observations for all sites in the **sites** dataset even if these sites are not observed on **bird_subset**.

```{r}
birds_full <- full_join(bird_subset,sites)
```

Next, an inner join. Note here that an `inner_join()` between **bird_subset** and **taxa_list** will contain only observations of `species_id` on both datasets.

```{r}
birds_inner <- inner_join(bird_subset, taxa_list, by = "species_id")
```

Let's suppose we used a `left_join()`. Note now that this keeps observations of birds with `species_id` that do not appear on the **taxa_list**.

```{r}
birds_inner_versus_left <- left_join(bird_subset, taxa_list, by = "species_id")
```

Now, let's try some data wrangling. First, rename `notes` to `bird_obs_notes` to avoid conflict with notes in the **surveys** dataset.

```{r}
birds <- bird_observations %>%
  rename(bird_obs_notes = notes)
```

Next, create a subset with all observations and join. Limit this subset to `survey_date`, `common_name`, `park_name`, and `bird_count`.

```{r}
birds_dataset <- birds %>%
  left_join(y = taxa_list, by = "species_id") %>%
  left_join(y = sites, by = "site_id") %>%
  left_join(y = surveys, by = c("survey_id", "site_id")) %>%
  select(survey_date, common_name, park_name, bird_count)
```

Use **lubridate** package to add a new column `survey_month`, which contains only the month number. And move this column.

```{r}
birds_1 <- birds_dataset %>%
  mutate(survey_month = as.factor(lubridate::month(survey_date))) %>%
  dplyr::relocate(survey_month, .after = survey_date)
```

Finally, let's find the total of birds observed by park and month and limit to "Lindo", "Orme", "Palomino" and "Sonrisa" parks.

```{r}
birds_clean <- birds_1 %>%
  group_by(park_name, survey_month) %>%
  summarize(sum(bird_count, na.rm = TRUE)) %>%
  filter(park_name %in% c("Lindo", "Orme", "Palomino", "Sonrisa"))
```

### Citation

Warren, P.S., A. Kinzig, C.A. Martin, and L. Machabee. 2013. Ecological and social Interactions in urban parks: bird surveys in local parks in the central Arizona-Phoenix metropolitan area ver 9. Environmental Data Initiative. <https://doi.org/10.6073/pasta/bff8e94e254c9317fa42c86f413a46ce> (Accessed 2022-08-30).
